---
title: "FinalCodeBirth"
output: html_document
date: "2026-04-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(ggplot2)
library(survival)
library(survminer)
library(dplyr)
library(gt)
library(car)
library(contsurvplot)
library(factoextra)
library(ggpubr)
library(RColorBrewer)
library(riskRegression)
library(pammtools)
library(gtsummary)
library(purrr)
library(cowplot)
library(phyloseq)
library(tidyverse)
library(dplyr)
library(DT)
library(decontam)
library(vegan)
library(scales)
library(lubridate)
library(microbiome)
library(gtsummary)
library(reshape2)
library(broom)
library(clusterCrit)
library(clusterSim)
library(ade4)


set.seed(1234)


```

MetaData and attach to phyloseq
```{r}

ps_birth <-readRDS("/Users/rapiduser/Desktop/HOPE_Manuscript/ps_birth.RDS")

birthmeta <- read_csv("/Users/rapiduser/Desktop/HOPE_Manuscript/16Sbirthmeta.csv")

ps_birth <- prune_taxa(!(taxa_sums(ps_birth) < 1), ps_birth)

birthmeta <- as.data.frame(birthmeta)

rownames(birthmeta) <- birthmeta$Subject

birthmeta_sampdata = sample_data(birthmeta)

sample_data(ps_birth) <- birthmeta_sampdata

milkbox <- read_csv("/Users/rapiduser/Desktop/HOPE_Manuscript/milkboxtransfer.csv")

```

Compute alpha diversity
```{r}

alpha_div <- estimate_richness(ps_birth, measures = c("Shannon", "Observed")) %>%
  rownames_to_column("named")

meta <- sample_data(ps_birth) %>%
  data.frame(.) %>%
  rownames_to_column("named")

meta$named <- gsub("-", ".", meta$named)

# Join together
alpha_div_meta <- left_join(alpha_div, meta, by = "named")

alpha_div_meta <- alpha_div_meta %>% 
  mutate(type_of_feeding_at_dischar = case_when(type_of_feeding_at_dischar == "Both"~"Breast Milk", 
                                                TRUE~type_of_feeding_at_dischar))
alpha_div_meta <- alpha_div_meta %>% 
  mutate(
    Mode.of.delivery = case_when(
      Mode.of.delivery == "Cesarean delivery" ~ "Cesarean",
      Mode.of.delivery == "Vaginal delivery" ~ "Vaginal",
      TRUE ~ Mode.of.delivery),
    Mode.of.delivery = factor(
      Mode.of.delivery,
      levels = c("Vaginal", "Cesarean")))

wilcox.test(Observed ~ Mode.of.delivery, data = alpha_div_meta, var.equal = TRUE)
wilcox.test(Shannon ~ Mode.of.delivery, data = alpha_div_meta, var.equal = TRUE)

wilcox.test(Observed ~ abxindicate, data = alpha_div_meta, var.equal = TRUE)
wilcox.test(Shannon ~ abxindicate, data = alpha_div_meta, var.equal = TRUE)

wilcox.test(Observed ~ type_of_feeding_at_dischar, data = alpha_div_meta)
wilcox.test(Shannon ~ type_of_feeding_at_dischar, data = alpha_div_meta)

alpha_div_meta %>% 
  filter(type_of_feeding_at_dischar == "Formula") %>% 
  summarise(
    q25 = quantile(Observed, 0.25, na.rm = TRUE),
    median = quantile(Observed, 0.5, na.rm = TRUE),
    q75 = quantile(Observed, 0.75, na.rm = TRUE))

alpha_div_meta %>% 
  filter(type_of_feeding_at_dischar == "Breast Milk") %>% 
  summarise(
    q25 = quantile(Observed, 0.25, na.rm = TRUE),
    median = quantile(Observed, 0.5, na.rm = TRUE),
    q75 = quantile(Observed, 0.75, na.rm = TRUE))

alpha_div_meta %>% 
  filter(type_of_feeding_at_dischar == "Formula") %>% 
  summarise(
    q25 = quantile(Shannon, 0.25, na.rm = TRUE),
    median = quantile(Shannon, 0.5, na.rm = TRUE),
    q75 = quantile(Shannon, 0.75, na.rm = TRUE))

alpha_div_meta %>% 
  filter(type_of_feeding_at_dischar == "Breast Milk") %>% 
  summarise(
    q25 = quantile(Shannon, 0.25, na.rm = TRUE),
    median = quantile(Shannon, 0.5, na.rm = TRUE),
    q75 = quantile(Shannon, 0.75, na.rm = TRUE))


```

Presence/Absence 
```{r}
ps_genus <- tax_glom(ps_birth, taxrank = "Genus", NArm = FALSE)

ps_relative <- transform_sample_counts(ps_genus, function(x) x/sum(x))

genus_df <- psmelt(ps_relative)
    
strep_abundance <- genus_df %>%
      filter(Genus %in% c("Streptococcus", "Staphylococcus", "Haemophilus", "Dolosigranulum", "Moraxella", "Corynebacterium")) %>%
      dplyr::select(Subject, Genus, Abundance)

strepabundOM <- merge(alpha_div_meta, strep_abundance, by = "Subject")

strepabundOM <- strepabundOM %>% 
  mutate(anyindicate = case_when(Abundance == 0~"No",
                                   Abundance > 0~"Yes"))

strepabundOM <- strepabundOM %>%
  mutate(anyindicate_num = ifelse(anyindicate == "Yes", 1, 0))

strepabundOM$abxindicate <- factor(
  strepabundOM$abxindicate,
  levels = c("No", "Yes"))

genera_keep <- c("Streptococcus","Staphylococcus","Haemophilus","Dolosigranulum","Moraxella", "Corynebacterium")

genus_counts <- strepabundOM %>%
  filter(Genus %in% genera_keep) %>%
  group_by(Genus, abxindicate) %>%
  summarise(
    successes = sum(anyindicate_num, na.rm = TRUE),
    total = n(),
    .groups = "drop") %>%
  mutate(prop = successes / total)


prop_results <- genus_counts %>%
  group_by(Genus) %>%
  summarise(
    diff_prop = prop[1] - prop[2],
    chisq = prop.test(x = successes, n = total, correct = FALSE)$statistic,
    p_value = prop.test(x = successes,n = total,correct = FALSE)$p.value,
    .groups = "drop")

final_results <- genus_counts %>%
  left_join(prop_results, by = "Genus")

final_results

#Switch to mode of delivery
genus_counts <- strepabundOM %>%
  filter(Genus %in% genera_keep) %>%
  group_by(Genus, Mode.of.delivery) %>%
  summarise(
    successes = sum(anyindicate_num, na.rm = TRUE),
    total = n(),
    .groups = "drop") %>%
  mutate(prop = successes / total)

prop_results <- genus_counts %>%
  group_by(Genus) %>%
  summarise(diff_prop = prop[1] - prop[2],
    chisq = prop.test(
      x = successes,
      n = total,
      correct = FALSE)$statistic,
    p_value = prop.test(
      x = successes,
      n = total,
      correct = FALSE)$p.value,
    conf_low = prop.test(
      x = successes,
      n = total,
      correct = FALSE)$conf.int[1],
    conf_high = prop.test(
      x = successes,
      n = total,
      correct = FALSE)$conf.int[2],
    .groups = "drop")

final_results <- genus_counts %>%
  left_join(prop_results, by = "Genus")

final_results

#Switch to Breastfeeding
genus_counts <- strepabundOM %>%
  filter(Genus %in% genera_keep) %>%
  group_by(Genus, type_of_feeding_at_dischar) %>%
  summarise(
    successes = sum(anyindicate_num, na.rm = TRUE),
    total = n(),
    .groups = "drop") %>%
  mutate(prop = successes / total)


prop_results <- genus_counts %>%
  group_by(Genus) %>%
  summarise(
    diff_prop = prop[1] - prop[2],
    chisq = prop.test(
      x = successes,
      n = total,
      correct = FALSE)$statistic,
    p_value = prop.test(
      x = successes,
      n = total,
      correct = FALSE)$p.value,
    conf_low = prop.test(
      x = successes,
      n = total,
      correct = FALSE)$conf.int[1],
    conf_high = prop.test(
      x = successes,
      n = total,
      correct = FALSE)$conf.int[2],
    .groups = "drop")

final_results <- genus_counts %>%
  left_join(prop_results, by = "Genus")

final_results

#Switch to Breastfeeding
genus_counts <- strepabundOM %>%
  filter(Genus %in% genera_keep) %>%
  group_by(Genus) %>%
  summarise(
    successes = sum(anyindicate_num, na.rm = TRUE),
    total = n(),
    .groups = "drop") %>%
  mutate(prop = successes / total)


prop_results <- genus_counts %>%
  group_by(Genus) %>%
  summarise(
    diff_prop = prop[1] - prop[2],
    chisq = prop.test(
      x = successes,
      n = total,
      correct = FALSE)$statistic,
    p_value = prop.test(
      x = successes,
      n = total,
      correct = FALSE)$p.value,
    conf_low = prop.test(
      x = successes,
      n = total,
      correct = FALSE)$conf.int[1],
    conf_high = prop.test(
      x = successes,
      n = total,
      correct = FALSE)$conf.int[2],
    .groups = "drop")

final_results <- genus_counts %>%
  left_join(prop_results, by = "Genus")

final_results

genus_counts <- strepabundOM %>%
  filter(Genus %in% genera_keep) %>%
  group_by(Genus, abxindicate) %>%
  summarise(
    successes = sum(anyindicate_num, na.rm = TRUE),
    total = n(),
    .groups = "drop") %>%
  mutate(prop = successes / total)

# Counts for Haem
streptab <- strepabundOM %>%
  filter(Genus == "Streptococcus") %>% 
  group_by(Mode.of.delivery) %>%
  summarise(
    successes = sum(anyindicate_num),
    total = n())

prop.test(
  x = streptab$successes,
  n = streptab$total,
  correct = FALSE)

prop.test(streptab$successes, streptab$total, correct = FALSE)

# Counts for each Staph
staphtab <- strepabundOM %>%
  filter(Genus == "Staphylococcus") %>% 
  group_by(Mode.of.delivery) %>%
  summarise(
    successes = sum(anyindicate_num),
    total = n())

prop.test(
  x = staphtab$successes,
  n = staphtab$total,
  correct = FALSE)

prop.test(staphtab$successes, staphtab$total, correct = FALSE)
```

Relative Abundance MoD Plot
```{r}
ps.NP.r = transform_sample_counts(ps_birth,function(x) x/sum(x))

sampleDF = sample_data(ps.NP.r) %>%
  data.frame(.)

ps.Genus = tax_glom(ps.NP.r, taxrank = "Genus") 

genusDF = as.data.frame(cbind(t(otu_table(ps.Genus, taxa_are_rows = FALSE)), tax_table(ps.Genus)))

genusDF$Kingdom = NULL
genusDF$Phylum=NULL
genusDF$Class=NULL
genusDF$Order=NULL
genusDF$Family=NULL
rownames(genusDF) = genusDF$Genus
genusDF$Genus=NULL
genusDF$Species = NULL
genusDF = mutate_all(genusDF, function(x) as.numeric(as.character(x)))

genusDF$C1 = rowMeans(genusDF)
genusDF$C1 <- genusDF$C1 / sum(genusDF$C1)

Top = genusDF$C1

psC1 = subset_samples(ps.Genus, sample_data(ps.Genus)$Mode.of.delivery == "Cesarean delivery")

genusDF = as.data.frame(cbind(t(otu_table(psC1, taxa_are_rows = FALSE)), tax_table(psC1)))

genusDF$Kingdom = NULL
genusDF$Phylum=NULL
genusDF$Class=NULL
genusDF$Order=NULL
genusDF$Family=NULL
rownames(genusDF) = genusDF$Genus
genusDF$Genus=NULL
genusDF$Species = NULL
genusDF = mutate_all(genusDF, function(x) as.numeric(as.character(x)))

genusDF$C1 = rowMeans(genusDF)
genusDF$C1 <- genusDF$C1 / sum(genusDF$C1)

AvgDF1 = genusDF$C1 / sum(genusDF$C1)

psC1 = subset_samples(ps.Genus, sample_data(ps.Genus)$Mode.of.delivery == "Vaginal delivery")

genusDF = as.data.frame(cbind(t(otu_table(psC1, taxa_are_rows = FALSE)), tax_table(psC1)))

genusDF$Kingdom = NULL
genusDF$Phylum=NULL
genusDF$Class=NULL
genusDF$Order=NULL
genusDF$Family=NULL
rownames(genusDF) = genusDF$Genus
genusDF$Genus=NULL
genusDF$Species = NULL
genusDF = mutate_all(genusDF, function(x) as.numeric(as.character(x)))

genusDF$C1 = rowMeans(genusDF)
genusDF$C1 <- genusDF$C1 / sum(genusDF$C1)

AvgDF2 = genusDF$C1 / sum(genusDF$C1)

AvgDF = cbind(Top, AvgDF1, AvgDF2) %>%
  data.frame(.)

rownames(AvgDF) = rownames(genusDF)

AvgDF$Avg = rowMeans(AvgDF)

fig = top_n(AvgDF, 30, Top)

fig <- fig %>% 
  dplyr::select(AvgDF1, AvgDF2, Avg)

Other = 1 - colSums(fig)

fig = rbind(fig, Other)

fig$tax = rownames(fig)

fig <- fig %>% 
  mutate(taxfix = case_when(tax == "31"~ "Other", 
                            tax == tax ~tax))

fig = dplyr::select(fig, -Avg) %>%
  dplyr::rename('Cesarean delivery' = AvgDF1,
                'Vaginal delivery' = AvgDF2)

f = melt(fig) 

mycolors <- c(
  "#1f78b4", "#33a02c", "#e31a1c", "#ff7f00", "#6a3d9a",
  "#b15928", "#a6cee3", "#b2df8a", "#fb9a99", "#fdbf6f",
  "#cab2d6", "#ffff99", "#8dd3c7", "#ffffb3", "#bebada",
  "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5",
  "#a6d854", "#bc80bd", "#ccebc5", "#ffed6f", "#66c2a5",
  "#fc8d62", "#8da0cb", "#e78ac3", "black", "#ffd92f",
  "gray")

f$taxfix <- factor(f$taxfix, levels = c(
  "Acinetobacter", "Anoxybacillus", "Bacillus", "Bacteroidetes_[G-6]",
  "Capnocytophaga", "Corynebacterium", "Cupriavidus", "Dolosigranulum",
  "Enterococcus", "Fastidiosipila", "Fusobacterium",
  "Gardnerella", "Haemophilus", "Kocuria", "Lactobacillus",
  "Leptotrichia", "Microbacterium", "Micrococcus", "Moraxella",
  "Neisseria", "Novosphingobium", "Paracoccus", "Pedobacter",
  "Porphyromonas", "Prevotella", "Pseudomonas", "Roseomonas",
  "Rothia", "Sphingomonas", "Staphylococcus", "Stenotrophomonas",
  "Streptococcus", "Veillonella",
  "Other"))

RAMoD=ggplot(f, aes(fill = taxfix, y=value, x=variable)) +
  geom_bar(position="stack", stat="identity", color = "black", size = 0.25) +
  scale_fill_manual(values = mycolors) +
  xlab("Mode of Delivery") +
  ylab("Mean Relative Abundance") +
  labs(fill = "Genera", title = "") +
  theme_minimal() +
  scale_y_continuous(limits = c(0,1), expand = c(0, 0))+
  theme(legend.key.size = unit(0.4, "cm"),
    legend.text     = element_text(size = 9),
    legend.title    = element_text(size = 9))
RAMoD

```

Relative Abundance ABX Plot
```{r}

ps.NP.r = transform_sample_counts(ps_birth,function(x) x/sum(x))

sampleDF = sample_data(ps.NP.r) %>%
  data.frame(.)

ps.Genus = tax_glom(ps.NP.r, taxrank = "Genus") 

genusDF = as.data.frame(cbind(t(otu_table(ps.Genus, taxa_are_rows = FALSE)), tax_table(ps.Genus)))

genusDF$Kingdom = NULL
genusDF$Phylum=NULL
genusDF$Class=NULL
genusDF$Order=NULL
genusDF$Family=NULL
rownames(genusDF) = genusDF$Genus
genusDF$Genus=NULL
genusDF$Species = NULL
genusDF = mutate_all(genusDF, function(x) as.numeric(as.character(x)))

genusDF$C1 = rowMeans(genusDF)
genusDF$C1 <- genusDF$C1 / sum(genusDF$C1)

Top = genusDF$C1

psC1 = subset_samples(ps.Genus, sample_data(ps.Genus)$abxindicate == "Yes")

genusDF = as.data.frame(cbind(t(otu_table(psC1, taxa_are_rows = FALSE)), tax_table(psC1)))

genusDF$Kingdom = NULL
genusDF$Phylum=NULL
genusDF$Class=NULL
genusDF$Order=NULL
genusDF$Family=NULL
rownames(genusDF) = genusDF$Genus
genusDF$Genus=NULL
genusDF$Species = NULL
genusDF = mutate_all(genusDF, function(x) as.numeric(as.character(x)))

genusDF$C1 = rowMeans(genusDF)
genusDF$C1 <- genusDF$C1 / sum(genusDF$C1)

AvgDF1 = genusDF$C1 / sum(genusDF$C1)

psC1 = subset_samples(ps.Genus, sample_data(ps.Genus)$abxindicate == "No")

genusDF = as.data.frame(cbind(t(otu_table(psC1, taxa_are_rows = FALSE)), tax_table(psC1)))

genusDF$Kingdom = NULL
genusDF$Phylum=NULL
genusDF$Class=NULL
genusDF$Order=NULL
genusDF$Family=NULL
rownames(genusDF) = genusDF$Genus
genusDF$Genus=NULL
genusDF$Species = NULL
genusDF = mutate_all(genusDF, function(x) as.numeric(as.character(x)))

genusDF$C1 = rowMeans(genusDF)
genusDF$C1 <- genusDF$C1 / sum(genusDF$C1)

AvgDF2 = genusDF$C1 / sum(genusDF$C1)

AvgDF = cbind(Top, AvgDF1, AvgDF2) %>%
  data.frame(.)

rownames(AvgDF) = rownames(genusDF)

AvgDF$Avg = rowMeans(AvgDF)

fig = top_n(AvgDF, 30, Top)

fig <- fig %>% 
  dplyr::select(AvgDF1, AvgDF2, Avg)

Other = 1 - colSums(fig)

fig = rbind(fig, Other)

fig$tax = rownames(fig)

fig <- fig %>% 
  mutate(taxfix = case_when(tax == "31"~ "Other", 
                            tax == tax ~tax))

fig = dplyr::select(fig, -Avg) %>%
  dplyr::rename('Yes' = AvgDF1,
                'No' = AvgDF2)

f = melt(fig) 

mycolors <- c(
  "#1f78b4", "#33a02c", "#e31a1c", "#ff7f00", "#6a3d9a",
  "#b15928", "#a6cee3", "#b2df8a", "#fb9a99", "#fdbf6f",
  "#cab2d6", "#ffff99", "#8dd3c7", "#ffffb3", "#bebada",
  "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5",
  "#a6d854", "#bc80bd", "#ccebc5", "#ffed6f", "#66c2a5",
  "#fc8d62", "#8da0cb", "#e78ac3", "black", "#ffd92f",
  "gray")

f$taxfix <- factor(f$taxfix, levels = c(
  "Acinetobacter", "Anoxybacillus", "Bacillus", "Bacteroidetes_[G-6]",
  "Capnocytophaga", "Corynebacterium", "Cupriavidus", "Dolosigranulum",
  "Enterococcus", "Fastidiosipila", "Fusobacterium",
  "Gardnerella", "Haemophilus", "Kocuria", "Lactobacillus",
  "Leptotrichia", "Microbacterium", "Micrococcus", "Moraxella",
  "Neisseria", "Novosphingobium", "Paracoccus", "Pedobacter",
  "Porphyromonas", "Prevotella", "Pseudomonas", "Roseomonas",
  "Rothia", "Sphingomonas", "Staphylococcus", "Stenotrophomonas",
  "Streptococcus", "Veillonella",
  "Other"))

RAABX=ggplot(f, aes(fill = taxfix, y=value, x=variable)) +
  geom_bar(position="stack", stat="identity", color = "black", size = 0.25) +
  scale_fill_manual(values = mycolors) +
  xlab("Intrapartum Antibiotics") +
  ylab("Mean Relative Abundance") +
  labs(fill = "Genera") +
  theme_minimal() +
  scale_y_continuous(limits = c(0,1), expand = c(0, 0))+
 theme(legend.key.size = unit(0.4, "cm"),
    legend.text     = element_text(size = 9),
    legend.title    = element_text(size = 9))

RAABX

```

Relative Abundance Feed Plot
```{r}

ps.NP.r = transform_sample_counts(ps_birth,function(x) x/sum(x))

sampleDF = sample_data(ps.NP.r) %>%
  data.frame(.)

ps.Genus = tax_glom(ps.NP.r, taxrank = "Genus") 

genusDF = as.data.frame(cbind(t(otu_table(ps.Genus, taxa_are_rows = FALSE)), tax_table(ps.Genus)))

genusDF$Kingdom = NULL
genusDF$Phylum=NULL
genusDF$Class=NULL
genusDF$Order=NULL
genusDF$Family=NULL
rownames(genusDF) = genusDF$Genus
genusDF$Genus=NULL
genusDF$Species = NULL
genusDF = mutate_all(genusDF, function(x) as.numeric(as.character(x)))

genusDF$C1 = rowMeans(genusDF)
genusDF$C1 <- genusDF$C1 / sum(genusDF$C1)

Top = genusDF$C1

psC1 = subset_samples(ps.Genus, sample_data(ps.Genus)$type_of_feeding_at_dischar == "Breast Milk")

genusDF = as.data.frame(cbind(t(otu_table(psC1, taxa_are_rows = FALSE)), tax_table(psC1)))

genusDF$Kingdom = NULL
genusDF$Phylum=NULL
genusDF$Class=NULL
genusDF$Order=NULL
genusDF$Family=NULL
rownames(genusDF) = genusDF$Genus
genusDF$Genus=NULL
genusDF$Species = NULL
genusDF = mutate_all(genusDF, function(x) as.numeric(as.character(x)))

genusDF$C1 = rowMeans(genusDF)
genusDF$C1 <- genusDF$C1 / sum(genusDF$C1)

AvgDF1 = genusDF$C1 / sum(genusDF$C1)

psC1 = subset_samples(ps.Genus, sample_data(ps.Genus)$type_of_feeding_at_dischar == "Formula")

genusDF = as.data.frame(cbind(t(otu_table(psC1, taxa_are_rows = FALSE)), tax_table(psC1)))

genusDF$Kingdom = NULL
genusDF$Phylum=NULL
genusDF$Class=NULL
genusDF$Order=NULL
genusDF$Family=NULL
rownames(genusDF) = genusDF$Genus
genusDF$Genus=NULL
genusDF$Species = NULL
genusDF = mutate_all(genusDF, function(x) as.numeric(as.character(x)))

genusDF$C1 = rowMeans(genusDF)
genusDF$C1 <- genusDF$C1 / sum(genusDF$C1)

AvgDF2 = genusDF$C1 / sum(genusDF$C1)

AvgDF = cbind(Top, AvgDF1, AvgDF2) %>%
  data.frame(.)

rownames(AvgDF) = rownames(genusDF)

AvgDF$Avg = rowMeans(AvgDF)

fig = top_n(AvgDF, 30, Top)

fig <- fig %>% 
  dplyr::select(AvgDF1, AvgDF2, Avg)

Other = 1 - colSums(fig)

fig = rbind(fig, Other)

fig$tax = rownames(fig)

fig <- fig %>% 
  mutate(taxfix = case_when(tax == "31"~ "Other", 
                            tax == tax ~tax))

fig = dplyr::select(fig, -Avg) %>%
  dplyr::rename('Breast Milk' = AvgDF1,
                'Formula' = AvgDF2)

f = melt(fig) 

mycolors <- c(
  "#1f78b4", "#33a02c", "#e31a1c", "#ff7f00", "#6a3d9a",
  "#b15928", "#a6cee3", "#b2df8a", "#fb9a99", "#fdbf6f",
  "#cab2d6", "#ffff99", "#8dd3c7", "#ffffb3", "#bebada",
  "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5",
  "#a6d854", "#bc80bd", "#ccebc5", "#ffed6f", "#66c2a5",
  "#fc8d62", "#8da0cb", "#e78ac3", "black", "#ffd92f",
  "gray")

f$taxfix <- factor(f$taxfix, levels = c(
  "Acinetobacter", "Anoxybacillus", "Bacillus", "Bacteroidetes_[G-6]",
  "Capnocytophaga", "Corynebacterium", "Cupriavidus", "Dolosigranulum",
  "Enterococcus", "Fastidiosipila", "Fusobacterium",
  "Gardnerella", "Haemophilus", "Kocuria", "Lactobacillus",
  "Leptotrichia", "Microbacterium", "Micrococcus", "Moraxella",
  "Neisseria", "Novosphingobium", "Paracoccus", "Pedobacter",
  "Porphyromonas", "Prevotella", "Pseudomonas", "Roseomonas",
  "Rothia", "Sphingomonas", "Staphylococcus", "Stenotrophomonas",
  "Streptococcus", "Veillonella",
  "Other"))

RAfeed=ggplot(f, aes(fill = taxfix, y=value, x=variable)) +
  geom_bar(position="stack", stat="identity", color = "black", size = 0.25) +
  scale_fill_manual(values = mycolors) +
  xlab("Feeding at Discharge") +
  ylab("Mean Relative Abundance") +
  labs(fill = "Genera") +
  theme_minimal() +
  scale_y_continuous(limits = c(0,1), expand = c(0, 0))+
 theme(legend.key.size = unit(0.4, "cm"),
    legend.text     = element_text(size = 9),
    legend.title    = element_text(size = 9))
RAfeed

RAfeed_mod <- RAfeed +
  guides(fill = guide_legend(ncol = 3, byrow = TRUE)) +
  theme(legend.position = "right")

```

```{r}
ps_genus <- tax_glom(ps_birth, taxrank = "Genus", NArm = FALSE)
seq.rel <- microbiome::transform(ps_genus, "compositional")
ASV <- abundances(pseq.rel)
meta <- meta(pseq.rel)

permanovaMOD <- adonis2(t(ASV) ~ Mode.of.delivery,
               data = meta, permutations=99, method = "bray")
permanovaMOD


permanovaABX <- adonis2(t(ASV) ~ abxindicate,
               data = meta, permutations=99, method = "bray")
permanovaABX


permanovaFEED <- adonis2(t(ASV) ~ type_of_feeding_at_dischar,
               data = meta, permutations=99, method = "bray")
permanovaFEED


```


Figure 1
```{r}

ShannnonMoD <- ggplot(alpha_div_meta, aes(x = Mode.of.delivery, y = Shannon)) +
  geom_violin(aes(fill = Mode.of.delivery),trim = TRUE) +
  geom_jitter(width = 0.15, alpha = 0.6, size = 1.8) +
  stat_summary(
    fun = median,
    geom = "crossbar",
    width = 0.35,
    linewidth = 0.9)+
  labs(x = "Mode of Delivery", y = "Shannon Diversity", fill = "Mode of Delivery")+
  theme_bw()+ theme(legend.key.size = unit(0.4, "cm"),
    legend.text = element_text(size = 9),
    legend.title = element_text(size = 9))

ObservedMoD <- ggplot(alpha_div_meta, aes(x = Mode.of.delivery, y = Observed)) +
  geom_violin(aes(fill = Mode.of.delivery),trim = TRUE) +
  geom_jitter(width = 0.15, alpha = 0.6, size = 1.8) +
  stat_summary(
    fun = median,
    geom = "crossbar",
    width = 0.35,
    linewidth = 0.9)+
  labs(x = "Mode of Delivery", y = "Richness", fill = "Mode of Delivery")+
  theme_bw()+ theme(legend.key.size = unit(0.4, "cm"),
    legend.text = element_text(size = 9),
    legend.title = element_text(size = 9))

ShannonABX <- ggplot(alpha_div_meta, aes(x = abxindicate, y = Shannon)) +
  geom_violin(aes(fill = abxindicate),trim = TRUE) +
  geom_jitter(width = 0.15, alpha = 0.6, size = 1.8) +
  stat_summary(
    fun = median,
    geom = "crossbar",
    width = 0.35,
    linewidth = 0.9)+
  labs(x = "Intrapartum Antibiotics", y = "Shannon Diversity", fill = "Intrapartum Antibiotics")+
  theme_bw()+ theme(legend.key.size = unit(0.4, "cm"),
    legend.text = element_text(size = 9),
    legend.title = element_text(size = 9))

ObservedABX <- ggplot(alpha_div_meta, aes(x = abxindicate, y = Observed)) +
  geom_violin(aes(fill = abxindicate), trim = TRUE) +
  geom_jitter(width = 0.15, alpha = 0.6, size = 1.8) +
  stat_summary(
    fun = median,
    geom = "crossbar",
    width = 0.35,
    linewidth = 0.9) +
  labs(x = "Intrapartum Antibiotics", y = "Richness", fill = "Intrapartum Antibiotics") +
  theme_bw() +
   theme(legend.key.size = unit(0.4, "cm"),
    legend.text  = element_text(size = 9),
    legend.title = element_text(size = 9))

ShannonFeed <- ggplot(alpha_div_meta, aes(x = type_of_feeding_at_dischar, y = Shannon)) +
  geom_violin(aes(fill = type_of_feeding_at_dischar),trim = TRUE) +
  geom_jitter(width = 0.15, alpha = 0.6, size = 1.8) +
  stat_summary(
    fun = median,
    geom = "crossbar",
    width = 0.35,
    linewidth = 0.9)+
  labs(x = "Feeding at Discharge", y = "Shannon Diversity", fill = "Feeding at Discharge")+
  theme_bw()+ theme(legend.key.size = unit(0.4, "cm"),
    legend.text = element_text(size = 9),
    legend.title = element_text(size = 9))

ObservedFeed <- ggplot(alpha_div_meta, aes(x = type_of_feeding_at_dischar, y = Observed)) +
  geom_violin(aes(fill = type_of_feeding_at_dischar),trim = TRUE) +
  geom_jitter(width = 0.15, alpha = 0.6, size = 1.8) +
  stat_summary(
    fun = median,
    geom = "crossbar",
    width = 0.35,
    linewidth = 0.9)+
  labs(x = "Feeding at Dischargee", y = "Richness", fill = "Feeding at Discharge")+
  theme_bw()+ theme(legend.key.size = unit(0.4, "cm"),
    legend.text = element_text(size = 9),
    legend.title = element_text(size = 9))

strep_summary <- strepabundOM %>% 
  mutate(anyindicate_num = ifelse(anyindicate == "Yes", 1, 0)) %>% 
  group_by(Genus, Mode.of.delivery) %>% 
  summarise(
    mean_strep = mean(anyindicate_num, na.rm = TRUE),
    se_strep   = sd(anyindicate_num, na.rm = TRUE) / sqrt(n()),
    .groups = "drop")

GDMoD <- ggplot(strep_summary, aes(x = Genus, y = mean_strep, fill = Mode.of.delivery)) +
  geom_col(position = position_dodge(width = 0.9)) +
 labs(title = "", x = "Genera", y = "Proportion", fill = "Mode of Delivery")+
  theme_bw() +
  theme(legend.key = element_rect(color = "black", size = 0.3),
    legend.key.size = unit(0.4, "cm"),
    legend.text = element_text(size = 9),
    legend.title = element_text(size = 9),
    axis.text.x = element_text(angle = 25, hjust = 1, vjust = 1))

strep_summary <- strepabundOM %>% 
  mutate(anyindicate_num = ifelse(anyindicate == "Yes", 1, 0)) %>% 
  group_by(Genus, abxindicate) %>% 
  summarise(
    mean_strep = mean(anyindicate_num, na.rm = TRUE),
    se_strep   = sd(anyindicate_num, na.rm = TRUE) / sqrt(n()),
    .groups = "drop")

GDABX <- ggplot(strep_summary, aes(x = Genus, y = mean_strep, fill = abxindicate)) +
  geom_col(position = position_dodge(width = 0.9)) +
  labs(title = "", x = "Genera",y = "Proportion", fill = "Intrapartum Antibiotics")+
  theme_bw() +
  theme(legend.key = element_rect(color = "black", size = 0.3),
    legend.key.size = unit(0.4, "cm"),
    legend.text = element_text(size = 9),
    legend.title = element_text(size = 9),
    axis.text.x = element_text(angle = 25, hjust = 1, vjust = 1))

strep_summary <- strepabundOM %>% 
  mutate(anyindicate_num = ifelse(anyindicate == "Yes", 1, 0)) %>% 
  group_by(Genus, type_of_feeding_at_dischar) %>% 
  summarise(
    mean_strep = mean(anyindicate_num, na.rm = TRUE),
    se_strep   = sd(anyindicate_num, na.rm = TRUE) / sqrt(n()),
    .groups = "drop")


GDFeed <- ggplot(strep_summary, aes(x = Genus, y = mean_strep, fill = type_of_feeding_at_dischar)) +
  geom_col(position = position_dodge(width = 0.9)) + 
  labs(title = "", x = "Genera", y = "Proportion", fill = "Feeding at Discharge")+
  theme_bw() +
  theme(legend.key = element_rect(color = "black", size = 0.3),
    legend.key.size = unit(0.4, "cm"),
    legend.text = element_text(size = 9),
    legend.title = element_text(size = 9),
    axis.text.x = element_text(angle = 25, hjust = 1, vjust = 1))

MoDFig <- plot_grid(
  plot_grid(
    ObservedMoD,
    ShannnonMoD,
    GDMoD,
    ncol = 1,
    labels = c("A", "B", "C"),
    align = "v"),
  plot_grid(NULL,RAMoD,NULL,
    ncol = 1,
    labels = c("", "D", ""),label_y = 1, label_x = -0.01,rel_heights = c(1, 3, 1)),
  ncol = 2,
  rel_widths = c(1, 1),
  align = "v")

ggsave(
  filename = "MoDFig.png",
  plot = MoDFig,
  width = 12,
  height = 10,
  dpi = 300)

ABXFig <- plot_grid(
  plot_grid(
    ObservedABX,
    ShannonABX,
    GDABX,
    ncol = 1,
    labels = c("A", "B", "C"),
    align = "v"),
  plot_grid(NULL,RAABX,NULL,
    ncol = 1,
    labels = c("", "D", ""),label_y = 1, label_x = -0.01,rel_heights = c(1, 3, 1)),
  ncol = 2,
  rel_widths = c(1, 1),
  align = "v")

ggsave(
  filename = "ABXFig.png",
  plot = ABXFig,
  width = 12,
  height = 10,
  dpi = 300)

feedFig <- plot_grid(
  plot_grid(
    ObservedFeed,
    ShannonFeed,
    GDFeed,
    ncol = 1,
    labels = c("A", "B", "C"),
    align = "v"),
  plot_grid(
    NULL,
    RAfeed,
    NULL,
    ncol = 1,
    labels = c("", "D", ""),
    label_y = 1,  
    label_x = -0.01,
    rel_heights = c(1, 3, 1)),
  ncol = 2,
  rel_widths = c(1, 1),
  align = "v")

ggsave(
  filename = "feedFig.png",
  plot = feedFig,
  width = 12,
  height = 10,
  dpi = 300, 
  bg = "white")

plot_grid(
  ObservedMoD,
  ShannnonMoD,
  GDMoD, 
  RAMoD,
  ObservedABX,
  ShannonABX,
  GDABX,
  RAABX,
  ObservedFeed,
  ShannonFeed,
  GDFeed,
  RAfeed,
  ncol = 4, labels = c("A","B","C","D","E","F", "G","H","I","J","K","L"), label_fontface = "bold")

nol <- theme(legend.position = "none")

ObservedMoD  <- ObservedMoD  + nol
ShannnonMoD  <- ShannnonMoD  + nol


ObservedABX  <- ObservedABX  + nol
ShannonABX   <- ShannonABX   + nol


ObservedFeed <- ObservedFeed + nol
ShannonFeed  <- ShannonFeed  + nol

topl <- theme(legend.position = "top")

GDMoD <- GDMoD + topl
GDABX  <- GDABX  + topl
GDFeed <- GDFeed + topl

GDMoD  <- GDMoD  + labs(fill = NULL, color = NULL)
GDABX  <- GDABX  + labs(fill = NULL, color = NULL)
GDFeed <- GDFeed + labs(fill = NULL, color = NULL)

nol <- theme(legend.position = "none")
RAMoD  <- RAMoD  + nol
RAABX  <- RAABX  + nol
RAfeed <- RAfeed + nol

RA_legend <- get_legend(
  RAMoD +
    theme(
      legend.position = "right",
      legend.key.height = unit(0.6, "cm"),
      legend.spacing.y = unit(0.5, "cm")))

RA_legend_plot <- ggdraw(RA_legend)

exposfigall <- plot_grid(plot_grid(
  ObservedMoD,  ShannnonMoD,  GDMoD,    RAMoD,
  ObservedABX,  ShannonABX,   GDABX,    RAABX,
  ObservedFeed, ShannonFeed,  GDFeed,   RAfeed,
  ncol = 4,
  labels = c("A","B","C","D","E","F","G","H","I", "J", "K", "L"),
  label_fontface = "bold"), plot_grid(
    NULL,
    RA_legend_plot,
    NULL,
    ncol = 1,
    labels = c("", "", ""),
    label_y = 1,  
    label_x = -0.01,
    rel_heights = c(0, 1, 0)),
  ncol = 2,
rel_heights = c(1, 2),
  rel_widths = c(1, 0.3),
  align = "v")

exposfigall

ggsave(
  filename = "exposefigall.png",
  plot = exposfigall,
  width = 17,
  height = 12,
  units = "in",
  dpi = 300, 
  limitsize = FALSE)

```

Figure 2 - Exposure cox models and forest plots
```{r}

cox_MoD <- coxph(Surv(time_2yr, OM_2yr) ~ Mode.of.delivery+RACEETH, data = alpha_div_meta)
sf_MoD <- survfit(Surv(time_2yr, OM_2yr) ~ Mode.of.delivery, data = alpha_div_meta)
p_MoD <- ggsurvplot(sf_MoD, data = alpha_div_meta,
                       xlab = "Age (days)",
                       ylab = "Fraction without AOM",
                       legend.title = "",
                       legend.labs = c("Vaginal", "Cesarean"),
                       title = "Mode of Delivery")$plot
p_MoD

cox_ABX <- coxph(Surv(time_2yr, OM_2yr) ~ abxindicate+RACEETH, data = alpha_div_meta)
sf_ABX <- survfit(Surv(time_2yr, OM_2yr) ~ abxindicate, data = alpha_div_meta)
p_ABX <- ggsurvplot(sf_ABX, data = alpha_div_meta,
                       xlab = "Age (days)",
                       ylab = "Fraction without AOM",
                       legend.title = "",
                       legend.labs = c("No", "Yes"),
                       title = "Intrapartum Antibiotics")$plot
p_ABX

milkbox <- milkbox %>% 
  mutate(AOMstatus = case_when(Any_AOM_bf_24mth == "Yes"~1,
                               Any_AOM_bf_24mth == "No"~0))

cox_feed <- coxph(Surv(timeto, AOMstatus) ~ mixBMin6 + Intra_anti_flag+MODclean+RACEETH+WC_cnt_bf_2yr+GA_WEEK_DAYS_NUM, data = milkbox)
sf_feed <- survfit(Surv(timeto, AOMstatus) ~ mixBMin6, data = milkbox)
p_feed <- ggsurvplot(sf_feed, data = milkbox,
                       xlab = "Age (days)",
                       ylab = "Fraction without AOM",
                       legend.title = "",
                       legend.labs = c("Breast Milk", "Formula"),
                       title = "Feeding Type")$plot
p_feed

cox_models <- list(
  "Cesarean Delivery" = cox_MoD,
  "Intrapartum Antibiotics" = cox_ABX,
  "Feeding Type" = cox_feed)


exposure_order <- c(
  "Cesarean Delivery",
  "Intrapartum Antibiotics",
  "Feeding Type")

forest_df <- map_dfr(names(cox_models), function(exposure) {
  tidy(cox_models[[exposure]]) %>%
    filter(term != "(Intercept)") %>%
    slice(1) %>%
    mutate(
      Exposure = exposure,
      HR = exp(estimate),
      HR_lower = exp(estimate - 1.96 * std.error),
      HR_upper = exp(estimate + 1.96 * std.error)) %>%
    dplyr::select(Exposure, HR, HR_lower, HR_upper)}) %>%
  mutate(Exposure = factor(Exposure, levels = rev(exposure_order)))

# Forest plot 
ExposForest <- ggplot(forest_df, aes(x = Exposure, y = HR, ymin = HR_lower, ymax = HR_upper)) +
  geom_pointrange(size = 1) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  coord_flip() +
  labs(y = "Hazard Ratio (95% CI)", x = "Exposure", title = "", color = "") +
  theme_bw(base_size = 14) +
  theme(panel.border = element_rect(color = "black", fill = NA, size = 1))

ggsave(
  filename = "ExposForest.png",
  plot = ExposForest,
  width = 8,
  height = 6,
  dpi = 300)

exposurefig <- plot_grid(
  p_MoD,
  p_ABX,
  p_feed,
  ExposForest,
  ncol = 2,
  labels = c("A","B","C","D"),
  label_fontface = "bold")
exposurefig

ggsave(
  filename = "exposurefig.png",
  plot = exposurefig,
  width = 12,
  height = 10,
  dpi = 300)

```

Alpha Diversity Survival Analysis
```{r}
cox_shannon <- coxph(Surv(time_2yr, OM_2yr) ~ Shannon+hrstonas, data = alpha_div_meta, x=TRUE)
cox_shannon
summary(cox_shannon)

p_shannon <-plot_surv_area(time="time_2yr",
               status="OM_2yr",
               variable="Shannon",
               data=alpha_div_meta,
               model=cox_shannon, 
               title = "Shannon Index", 
               xlab = "Age (days)")+
  ylim(c(0, 1))

cox_observed <- coxph(Surv(time_2yr, OM_2yr) ~ Observed+hrstonas, data = alpha_div_meta, x=TRUE)
cox_observed

p_observed <-plot_surv_area(time="time_2yr",
               status="OM_2yr",
               variable="Observed",
               data=alpha_div_meta,
               model=cox_observed, 
               title = "Richness", 
               xlab = "Age (days)")+
  ylim(c(0, 1))
```

Presence/Absence Survival Analysis
```{r}

### Streptococcus -------------------------------------------------
df_strep <- strepabundOM %>%
  filter(Genus == "Streptococcus") %>%
  filter(complete.cases(time_2yr, OM_2yr, anyindicate))

cox_strep <- coxph(Surv(time_2yr, OM_2yr) ~ anyindicate, data = df_strep)
sf_strep <- survfit(Surv(time_2yr, OM_2yr) ~ anyindicate, data = df_strep)
p_strep <- ggsurvplot(sf_strep, data = df_strep,
                       xlab = "Age (days)",
                       ylab = "Fraction without AOM",
                       legend.title = "Presence",
                       legend.labs = c("No", "Yes"),
                       title = "Streptococcus")$plot+
  theme(plot.title = element_text(face = "italic"))

### Staphylococcus -----------------------------------------------
df_staph <- strepabundOM %>%
  filter(Genus == "Staphylococcus") %>%
  filter(complete.cases(time_2yr, OM_2yr, anyindicate))

cox_staph <- coxph(Surv(time_2yr, OM_2yr) ~ anyindicate, data = df_staph)
sf_staph <- survfit(Surv(time_2yr, OM_2yr) ~ anyindicate, data = df_staph)
p_staph <- ggsurvplot(sf_staph, data = df_staph,
                       xlab = "Age (days)",
                       ylab = "Fraction without AOM",
                       legend.title = "Presence",
                       legend.labs = c("No", "Yes"),
                       title = "Staphylococcus")$plot+
  theme(plot.title = element_text(face = "italic"))

### Haemophilus ---------------------------------------------------
df_haem <- strepabundOM %>%
  filter(Genus == "Haemophilus") %>%
  filter(complete.cases(time_2yr, OM_2yr, anyindicate))

cox_haem <- coxph(Surv(time_2yr, OM_2yr) ~ anyindicate, data = df_haem)
sf_haem <- survfit(Surv(time_2yr, OM_2yr) ~ anyindicate, data = df_haem)
p_haem <- ggsurvplot(sf_haem, data = df_haem,
                       xlab = "Age (days)",
                       ylab = "Fraction without AOM",
                       legend.title = "Presence",
                       legend.labs = c("No", "Yes"),
                       title = "Haemophilus")$plot+
  theme(plot.title = element_text(face = "italic"))

### Dolosigranulum -----------------------------------------------
df_dolo <- strepabundOM %>%
  filter(Genus == "Dolosigranulum") %>%
  filter(complete.cases(time_2yr, OM_2yr, anyindicate))

cox_dolo <- coxph(Surv(time_2yr, OM_2yr) ~ anyindicate, data = df_dolo)
sf_dolo <- survfit(Surv(time_2yr, OM_2yr) ~ anyindicate, data = df_dolo)
p_dolo <- ggsurvplot(sf_dolo, data = df_dolo,
                       xlab = "Age (days)",
                       ylab = "Fraction without AOM",
                       legend.title = "Presence",
                       legend.labs = c("No", "Yes"),
                       title = "Dolosigranulum")$plot+
  theme(plot.title = element_text(face = "italic"))

### Moraxella -----------------------------------------------------
df_morax <- strepabundOM %>%
  filter(Genus == "Moraxella") %>%
  filter(complete.cases(time_2yr, OM_2yr, anyindicate))

cox_morax <- coxph(Surv(time_2yr, OM_2yr) ~ anyindicate, data = df_morax)
sf_morax <- survfit(Surv(time_2yr, OM_2yr) ~ anyindicate, data = df_morax)
p_morax <- ggsurvplot(sf_morax, data = df_morax,
                       xlab = "Age (days)",
                       ylab = "Fraction without AOM",
                       legend.title = "Presence",
                       legend.labs = c("No", "Yes"),
                       title = "Moraxella")$plot+
  theme(plot.title = element_text(face = "italic"))

### Corynebacterium ----------------------------------------------
df_cory <- strepabundOM %>%
  filter(Genus == "Corynebacterium") %>%
  filter(complete.cases(time_2yr, OM_2yr, anyindicate))

cox_cory <- coxph(Surv(time_2yr, OM_2yr) ~ anyindicate, data = df_cory)
sf_cory <- survfit(Surv(time_2yr, OM_2yr) ~ anyindicate, data = df_cory)
p_cory <- ggsurvplot(sf_cory, data = df_cory,
                       xlab = "Age (days)",
                       ylab = "Fraction without AOM",
                       legend.title = "Presence",
                       legend.labs = c("No", "Yes"),
                       title = "Corynebacterium")$plot+
  theme(plot.title = element_text(face = "italic"))

```

Figure 3 - Microbiome Survival
```{r}
cox_models <- list(
  "Shannon Diversity" = cox_shannon,
  "Richness" = cox_observed,
  "Streptococcus" = cox_strep,
  "Staphylococcus" = cox_staph,
  "Haemophilus" = cox_haem,
  "Dolosigranulum" = cox_dolo,
  "Moraxella" = cox_morax,
  "Corynebacterium" = cox_cory)

exposure_order <- c(
  "Shannon Diversity",
  "Richness",
  "Streptococcus",
  "Staphylococcus",
  "Haemophilus",
  "Dolosigranulum",
  "Moraxella",
  "Corynebacterium")

forest_df <- imap_dfr(cox_models, function(model, exposure) {
  td <- tidy(model) %>%
    filter(term != "(Intercept)") 
  td[1, ] %>%
    mutate(
      Exposure = exposure,
      HR = exp(estimate),
      HR_lower = exp(estimate - 1.96 * std.error),
      HR_upper = exp(estimate + 1.96 * std.error)) %>%
    dplyr::select(Exposure, HR, HR_lower, HR_upper)}) %>%
  mutate(Exposure = factor(Exposure, levels = rev(exposure_order)))

forestlabels <- rev(c(
  "Shannon Diversity",
  "Richness",
  expression(italic("Streptococcus")),
  expression(italic("Staphylococcus")),
  expression(italic("Haemophilus")),
  expression(italic("Dolosigranulum")),
  expression(italic("Moraxella")),
  expression(italic("Corynebacterium"))))

# Forest plot
Microbe_Forest <- ggplot(
  forest_df,
  aes(x = Exposure, y = HR, ymin = HR_lower, ymax = HR_upper)) +
  geom_pointrange(size = 1) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  coord_flip() +
  scale_x_discrete(labels = forestlabels) +
  labs(x = "Microbiome Features", y = "Hazard Ratio (95% CI)", title = "") +
  theme_bw(base_size = 14) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.grid.major.y = element_blank(),
    legend.position = "none")

Microbe_Forest

ggsave(
  filename = "Microbe_Forest.png",
  plot = Microbe_Forest,
  width = 8,
  height = 6,
  dpi = 300)

Microb_surv_fig <- plot_grid(
  p_shannon, 
  p_observed, 
  p_strep,
  p_staph,
  p_haem,
  p_dolo,
  p_morax,
  p_cory,
  Microbe_Forest,
  ncol = 3,
  labels = c("A","B","C","D","E","F", "G","H", "I"),
  label_fontface = "bold")

Microb_surv_fig

ggsave(
  filename = "Microb_surv_fig.png",
  plot = Microb_surv_fig,
  width = 12,
  height = 10,
  dpi = 300)

```

Cluster generation
```{r}

tax_table(ps_birth)

vegan_otu <- function(physeq) {
  OTU <- otu_table(physeq)
  if (taxa_are_rows(OTU)) {
    OTU <- t(OTU)}
  return(as(OTU, "matrix"))}

ps.Genus= tax_glom(ps_birth, taxrank = "Genus") %>%
  transform_sample_counts(.,function(x) x/sum(x)) %>%
  subset_taxa(., Genus!= "NA") 

DistType = t(as.data.frame(vegan_otu(ps.Genus)))

data.dist = vegdist(vegan_otu(ps.Genus), method = "bray")

pam.clustering=function(x,k) { # x is a distance matrix and k the number of clusters
  require(cluster)
  cluster = as.vector(pam(as.dist(x), k, diss=TRUE)$clustering)
  return(cluster)}

nclusters=NULL

for (k in 1:20) { 
  if (k==1) {
    nclusters[k]=NA } else {
    data.cluster_temp=pam.clustering(data.dist, k)
    nclusters[k]=index.G1(t(DistType), data.cluster_temp,  d = data.dist,
                          centrotypes = "medoids")}}

plot(nclusters, type="h", xlab="k clusters", ylab="CH index")

data.cluster=pam.clustering(data.dist, k=2)

nclusters = index.G1(t(DistType), data.cluster, d = data.dist, centrotypes = "medoids")

#look at number to see if it is highest average (or could change mean to median)
obs.silhouette=mean(silhouette(data.cluster, data.dist)[,3])

obs.pcoa=dudi.pco(data.dist, scannf=F, nf=2)

#PCOA of clusters
s.class(obs.pcoa$li, fac=as.factor(data.cluster), grid=T,  col = c(1:4))

p = cbind(obs.pcoa$li,data.cluster)

p$data.cluster = as.factor(p$data.cluster)

clusterspace=ggplot(data = p, aes(x = A1, y = A2)) +
  geom_point(data = p, aes(color = data.cluster), alpha=0.7) + 
  #facet_wrap(~ facet, ncol=4) +
  scale_color_manual(values = c("#4A72A6","#48A462","#7E6E85","#EC83BA","#FFB716","#e6550d","#5ab4ac")) + 
  theme_bw() +
  stat_ellipse(aes(color = data.cluster)) +
  stat_stars(aes(color = data.cluster), show.legend = T) +
  xlab("PC1 (8.5%)") + 
  ylab("PC2 (6.7%)") +
  labs(color = "Cluster")+ 
  scale_y_continuous(labels = scales::number_format(accuracy = 0.1))
clusterspace

max_k <- min(10, nsamples(ps_birth) - 1)
ch_scores <- numeric(max_k - 1)

optimal_k <- which.max(ch_scores) + 1

# PAM clustering
pam_final <- pam(data.dist, k = optimal_k, diss = TRUE)

# Silhouette plot
p3 <- fviz_silhouette(pam_final, print.summary = FALSE) +
  labs(title = "Silhouette Plot - Cluster Quality Assessment",
       subtitle = "Based on Genus-Species Level") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 11))

print(p3)

sample_data(ps.Genus)$cluster = as.factor(data.cluster)

sampleDF = sample_data(ps.Genus) %>%
  data.frame(.)

sampleDF <- sampleDF %>% 
 dplyr::select(Record, cluster)

sampleDF <- sampleDF[!duplicated(sampleDF$Record), ]

```

Cluster Relative Abundance
```{r}

sampleDF <- sample_data(sampleDF)

ps_birth <- merge_phyloseq(ps_birth, sampleDF)

ps.NP.r = subset_samples(ps_birth, sample_data(ps_birth)$Subject %in% alpha_div_meta$Subject)

ps.NP.r = transform_sample_counts(ps.NP.r,function(x) x/sum(x))

sampleDF = sample_data(ps.NP.r) %>%
  data.frame(.)

ps.Genus = tax_glom(ps.NP.r, taxrank = "Genus") 

genusDF = as.data.frame(cbind(t(otu_table(ps.Genus, taxa_are_rows = FALSE)), tax_table(ps.Genus)))

genusDF$Kingdom = NULL
genusDF$Phylum=NULL
genusDF$Class=NULL
genusDF$Order=NULL
genusDF$Family=NULL
rownames(genusDF) = genusDF$Genus
genusDF$Genus=NULL
genusDF$Species = NULL
genusDF = mutate_all(genusDF, function(x) as.numeric(as.character(x)))

genusDF$C1 = rowMeans(genusDF)
genusDF$C1 <- genusDF$C1 / sum(genusDF$C1)

Top = genusDF$C1

psC1 = subset_samples(ps.Genus, sample_data(ps.Genus)$cluster == 1)

genusDF = as.data.frame(cbind(t(otu_table(psC1, taxa_are_rows = FALSE)), tax_table(psC1)))

genusDF$Kingdom = NULL
genusDF$Phylum=NULL
genusDF$Class=NULL
genusDF$Order=NULL
genusDF$Family=NULL
rownames(genusDF) = genusDF$Genus
genusDF$Genus=NULL
genusDF$Species = NULL
genusDF = mutate_all(genusDF, function(x) as.numeric(as.character(x)))

genusDF$C1 = rowMeans(genusDF)

AvgDF1 = genusDF$C1 / sum(genusDF$C1)

psC1 = subset_samples(ps.Genus, sample_data(ps.Genus)$cluster == 2)

genusDF = as.data.frame(cbind(t(otu_table(psC1, taxa_are_rows = FALSE)), tax_table(psC1)))

genusDF$Kingdom = NULL
genusDF$Phylum=NULL
genusDF$Class=NULL
genusDF$Order=NULL
genusDF$Family=NULL
rownames(genusDF) = genusDF$Genus
genusDF$Genus=NULL
genusDF$Species = NULL
genusDF = mutate_all(genusDF, function(x) as.numeric(as.character(x)))

genusDF$C1 = rowMeans(genusDF)

AvgDF2 = genusDF$C1 / sum(genusDF$C1)

AvgDF = cbind(Top, AvgDF1, AvgDF2 ) %>%
  data.frame(.)

rownames(AvgDF) = rownames(genusDF)

AvgDF$Avg = rowMeans(AvgDF)

fig = top_n(AvgDF, 30, Top)

fig <- fig %>% 
  dplyr::select(AvgDF1, AvgDF2, Avg)

Other = sum(AvgDF$Avg) - colSums(fig)

fig = rbind(fig, Other)

fig$tax = rownames(fig)

fig <- fig %>% 
  mutate(taxfix = case_when(tax == "31"~ "Other", 
                            tax == tax ~tax))

fig = dplyr::select(fig, -Avg) %>%
  dplyr::rename('1' = AvgDF1,
                '2' = AvgDF2)

f = melt(fig) 

mycolors <- c(
  "#1f78b4", "#33a02c", "#e31a1c", "#ff7f00", "#6a3d9a",
  "#b15928", "#a6cee3", "#b2df8a", "#fb9a99", "#fdbf6f",
  "#cab2d6", "#ffff99", "#8dd3c7", "#ffffb3", "#bebada",
  "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5",
  "#a6d854", "#bc80bd", "#ccebc5", "#ffed6f", "#66c2a5",
  "#fc8d62", "#8da0cb", "#e78ac3", "black", "#ffd92f",
  "gray" 
)

f$taxfix <- factor(f$taxfix, levels = c(
  "Acinetobacter", "Anoxybacillus", "Bacillus", "Bacteroidetes_[G-6]",
  "Capnocytophaga", "Corynebacterium", "Cupriavidus", "Dolosigranulum",
  "Enterococcus", "Fastidiosipila", "Fusobacterium",
  "Gardnerella", "Haemophilus", "Kocuria", "Lactobacillus",
  "Leptotrichia", "Microbacterium", "Micrococcus", "Moraxella",
  "Neisseria", "Novosphingobium", "Paracoccus", "Pedobacter",
  "Porphyromonas", "Prevotella", "Pseudomonas", "Roseomonas",
  "Rothia", "Sphingomonas", "Staphylococcus", "Stenotrophomonas",
  "Streptococcus", "Veillonella",
  "Other"
))

RAcluster=ggplot(f, aes(fill = taxfix, y=value, x=variable)) +
  geom_bar(position="stack", stat="identity", color = "black", size = 0.25) +
  scale_fill_manual(values = mycolors) +
  xlab("Cluster") +
  ylab("Mean Relative Abundance") +
  labs(fill = "Genera", title = "") +
  theme_minimal() +
  scale_y_continuous(limits = c(0,1), expand = c(0, 0))+
theme(legend.text.align = 0)
RAcluster
```

Cluster Survival Analysis
```{r}

birthmeta <- sample_data(ps_birth) %>% 
  data.frame(.)

birthmeta$cluster <- as.character(birthmeta$cluster)

res.cox <- coxph(Surv(time_2yr, OM_2yr) ~ cluster+Mode.of.delivery+gacalcdel+hrstonas+ICUflag+hrsofrup, data = birthmeta)
res.cox

fit <- survfit(Surv(time_2yr, OM_2yr) ~ cluster, data = birthmeta)

coxcluster <- ggsurvplot(fit, data = birthmeta, ylab = "Fraction without AOM", xlab = "Age (days)", legend.labs = c("1", "2"), legend.title = "Cluster", legend = "right")

coxcluster$plot <- coxcluster$plot +
  scale_color_manual(values = c("#4A72A6","#48A462","#7E6E85","#EC83BA"))

coxcluster

birthmeta$cluster <- factor(birthmeta$cluster)
levels(birthmeta$cluster)

contrasts(birthmeta$cluster) <- contr.sum(length(levels(birthmeta$cluster)))

fit <- coxph(Surv(time_2yr, OM_2yr) ~ cluster, data = birthmeta)
summary(fit)
```

Figure X - Clusters
```{r}
bigfig <- plot_grid(clusterspace,coxcluster$plot, RAcluster,  labels = c("A", "B", "C"))
bigfig

clusterRA_cox <- plot_grid(
  bigfig,
  ncol = 1,
  rel_heights = c(1, 0.08))

clusterfig <- plot_grid(
  plot_grid(
    clusterspace,
    coxcluster$plot,
    ncol = 1,
    labels = c("A", "B"),
    align = "v"),
  plot_grid(
    NULL,
    RAcluster,
    NULL,
    ncol = 1,
    labels = c("", "C", ""),
    label_y = 1,  
    label_x = -0.01,
    rel_heights = c(1, 3, 1)),
  ncol = 2,
  rel_widths = c(1, 1),
  align = "v")
clusterfig

RAcluster_mod <- RAcluster +
  guides(fill = guide_legend(ncol = 4, byrow = TRUE)) +
  theme(legend.position = "right") 

legend <- get_legend(RAcluster_mod)

RAcluster_clean <- RAcluster + theme(legend.position = "none")

clusterfig <- plot_grid(
  
  plot_grid(
    clusterspace,
    coxcluster$plot,
    ncol = 2,
    labels = c("A", "B")),
  plot_grid(
    RAcluster_clean,
    legend,
    ncol = 2,
    labels = c("C", ""),
    rel_widths = c(1, 1)),
  ncol = 1,
  rel_heights = c(1, 1))

clusterfig

ggsave(
  filename = "clusterfig.png",
  plot = clusterfig,
  width = 12,
  height = 9,
  dpi = 300)
```

```{r}
ps_genus <- tax_glom(ps_birth, taxrank = "Genus", NArm = FALSE)
pseq.rel <- microbiome::transform(ps_genus, "compositional")
ASV <- abundances(pseq.rel)
meta <- meta(pseq.rel)

permanovaCLUST <- adonis2(t(ASV) ~ cluster,
               data = meta, permutations=99, method = "bray")
permanovaCLUST

```


Save Phyloseq
```{r}

saveRDS(ps_birth, "/Users/rapiduser/Desktop/HOPE_Manuscript/ps_birth.RDS")

```


```{r}

data <- sample_data(ps_birth) %>% 
  data.frame(.)


data %>% 
  dplyr::select(cluster, Mode.of.delivery, type_of_feeding_at_dischar, abxindicate, hrstonas, gacalcdel, ICUflag, gbsstatus) %>% 
  tbl_summary(by = cluster) %>% 
  add_p()

write.csv(data, "psbirthdata.csv")

```

Individual Microbiome Profiles
```{r}

tax_rank <- "Genus"  

physeq_genus <- tax_glom(ps_birth, taxrank = tax_rank)

physeq_genus_rel <- transform_sample_counts(physeq_genus, function(x) x / sum(x))

df <- psmelt(physeq_genus_rel)

top_genera <- c(
  "Acinetobacter", "Anoxybacillus", "Bacillus", "Bacteroidetes_[G-6]",
  "Capnocytophaga", "Corynebacterium", "Cupriavidus", "Dolosigranulum",
  "Enterococcus", "Fastidiosipila", "Fusobacterium",
  "Gardnerella", "Haemophilus", "Kocuria", "Lactobacillus",
  "Leptotrichia", "Microbacterium", "Micrococcus", "Moraxella",
  "Neisseria", "Novosphingobium", "Paracoccus", "Pedobacter",
  "Porphyromonas", "Prevotella", "Pseudomonas", "Roseomonas",
  "Rothia", "Sphingomonas", "Staphylococcus", "Stenotrophomonas",
  "Streptococcus", "Veillonella",
  "Other")

df$Genus2 <- ifelse(df$Genus %in% top_genera, as.character(df$Genus), "Other")

df$Genus2 <- factor(df$Genus2,levels = c(
  "Acinetobacter", "Anoxybacillus", "Bacillus", "Bacteroidetes_[G-6]",
  "Capnocytophaga", "Corynebacterium", "Cupriavidus", "Dolosigranulum",
  "Enterococcus", "Fastidiosipila", "Fusobacterium",
  "Gardnerella", "Haemophilus", "Kocuria", "Lactobacillus",
  "Leptotrichia", "Microbacterium", "Micrococcus", "Moraxella",
  "Neisseria", "Novosphingobium", "Paracoccus", "Pedobacter",
  "Porphyromonas", "Prevotella", "Pseudomonas", "Roseomonas",
  "Rothia", "Sphingomonas", "Staphylococcus", "Stenotrophomonas",
  "Streptococcus", "Veillonella",
  "Other"))

set.seed(123)  


genus_colors <- c(
  "#1f78b4", "#33a02c", "#e31a1c", "#ff7f00", "#6a3d9a",
  "#b15928", "#a6cee3", "#b2df8a", "#fb9a99", "#fdbf6f",
  "#cab2d6", "tan", "#8dd3c7", "#ffffb3", "#bebada",
  "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5",
  "#a6d854", "#bc80bd", "#ccebc5", "#ffed6f", "#66c2a5",
  "#fc8d62", "#8da0cb", "#e78ac3", "black", "#ffd92f",
  "gray" )

df$hrstonas <- round(df$hrstonas, digits = 1)
df <- df[!is.na(df$hrstonas),]
df <- df %>% 
  filter(hrstonas > 0)


CnoA <- df %>% 
  filter(Mode.of.delivery == "Cesarean delivery" & abxindicate == "No") %>%
  ggplot(aes(x = Sample, y = Abundance, fill = Genus2)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = genus_colors) +
  facet_wrap(~ hrstonas, nrow = 3, scales = "free_x") +
  theme_bw() +
  theme(
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank(),
    strip.text   = element_text(size = 8)) +
  labs(y = "Proportional Abundance",fill = "Genus (Top 30)", x = "Age (hours)", title = "Cesarean delivery and No Intrapartum Antibiotics")

VnoA <- df %>% 
  filter(Mode.of.delivery == "Vaginal delivery" & abxindicate == "No") %>%
  ggplot(aes(x = Sample, y = Abundance, fill = Genus2)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = genus_colors) +
  facet_wrap(~ hrstonas, nrow = 4, scales = "free_x") +
  theme_bw() +
  theme(
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank(),
    strip.text   = element_text(size = 8)) +
  labs(y = "Proportional Abundance",fill = "Genus (Top 30)", x = "Age (hours)", title = "Vaginal delivery and No Intrapartum Antibiotics")

CyesA <- df %>% 
  filter(Mode.of.delivery == "Cesarean delivery" & abxindicate == "Yes") %>%
  ggplot(aes(x = Sample, y = Abundance, fill = Genus2)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = genus_colors) +
  facet_wrap(~ hrstonas, nrow = 2, scales = "free_x") +
  theme_bw() +
  theme(
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank(),
    strip.text   = element_text(size = 8)) +
  labs(y = "Proportional Abundance",fill = "Genus (Top 30)", x = "Age (hours)", title = "Cesarean delivery and Intrapartum Antibiotics Administered")

VyesA <- df %>% 
  filter(Mode.of.delivery == "Vaginal delivery" & abxindicate == "Yes") %>%
  ggplot(aes(x = Sample, y = Abundance, fill = Genus2)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = genus_colors) +
  facet_wrap(~ hrstonas, nrow = 3, scales = "free_x") +
  theme_bw() +
  theme(
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank(),
    strip.text   = element_text(size = 8)) +
  labs(y = "Proportional Abundance", fill = "Genus (Top 30)", 
    x = "Age (hours)", 
    title = "Vaginal delivery and Intrapartum Antibiotics Administered")

ggsave(
  filename = "CnoA.png",
  plot = CnoA,
  width = 12,
  height = 8,
  dpi = 300)

ggsave(
  filename = "CyesA.png",
  plot = CyesA,
  width = 11,
  height = 6,
  dpi = 300)

ggsave(
  filename = "VnoA.png",
  plot = VnoA,
  width = 10,
  height = 8,
  dpi = 300)

ggsave(
  filename = "VyesA.png",
  plot = VyesA,
  width = 12,
  height = 7,
  dpi = 300)

```

